---
title: "functions-lesson"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Roxygen provides functions for parsing documentation commands.

```{r}
library(roxygen2)
```

# Example without using a function


```{r}
airtemps <- c(212, 30.3, 78, 32)
celsius1 <- (airtemps[1] - 32)*5/9
celsius2 <- (airtemps[2] - 32)*5/9
celsius3 <- (airtemps[3] - 32)*5/9

(airtemps - 32)*5/9

```

# Create a function

Because of function scoping, there is a new celsius variable within the function called celsius that can have a different value. Because it is inside the function, it is not visible outside the function. 

```{r}

fahr_to_celsius <- function(fahr) {
  celsius <- (fahr - 32)*5/9
  return(celsius)
}

```

# Test the function

```{r}
celsius <- fahr_to_celsius(airtemps[1])
celsius
celsius1 == celsius

```

```{r}
celsius <- fahr_to_celsius(airtemps)
celsius
```

# My turn to create a function

```{r}
#' Converts Celsius to Fahrenheit
#' 
#' This function takes scalar or vector values in degress Celsius
#' and returns values in degrees Fahrenheit
#'
#' @param celsius the input temperature in degrees Celsius

celsius_to_fahr <- function(celsius){
  fahr <- (celsius * 9/5) + 32
  return(fahr)
}
```

```{r}
fahr <- celsius_to_fahr(celsius)
fahr

fahr == airtemps
```




# Document the function